---
title: "R Notebook"
output: html_notebook
---

```{r}
data = read.csv("D:/Fish.csv")
View(data)
```

```{r}
head(data)
```

```{r}
tail(data)
```

```{r}
summary(data)
```

```{r}
str(data)
```

```{r}
dim(data)
```

```{r}
colSums(is.na(data))
```

```{r}
colnames(data)
```

```{r}
d<-na.omit(data)
colSums(is.na(d))
```


EDA PART:

```{r}
library(ggcorrplot)
corr=cor(d[,2:7])
ggcorrplot(corr, lab = TRUE)
```

```{r}
ggplot(data = d)+
  geom_bar(mapping = aes(y = d$Weight, x = d$ï..Species, fill = d$ï..Species), stat= "identity") +
  labs(title="weight per Species", x="Species", y="Weight in Grams")
```

```{r}
ggplot(data = d)+
  geom_point(mapping = aes(y = d$Weight, x = d$Length1, col = d$ï..Species), stat= "identity") +
  labs(title="Weight per Species", x="Length", y="Weight in Grams")
```

```{r}
ggplot(data = d,mapping = aes(y = d$Weight,col=d$ï..Species), stat= "identity")+
  geom_boxplot() +
  labs(title="weight per Species", x="Species", y="Weight in Grams")
```

```{r}
ggplot(d,aes(x=d$ï..Species,y=d$Weight,fill=d$ï..Species))+geom_boxplot()
```

```{r}
ggplot(d,aes(x=d$ï..Species,y=d$Length1,fill=d$ï..Species))+geom_boxplot()
```

```{r}
ggplot(d,aes(x=d$ï..Species,y=d$Height,fill=d$ï..Species))+geom_boxplot()
```

From the boxplots above, we can see that for Roach, there are at least one outlier for all the variables.
To further check the outliers, we can check the quantile.
Quantile for species Roach, in Weight:

```{r}
library(dplyr)
```

```{r}
d_Roach <- d %>% filter(d$ï..Species == 'Roach')
quantile(d_Roach$Weight)

```

From the quantile, we can tell that the last quantile is far from the rest of the quantiles, so this is the outlier situation.

To deal with this situation, we removed the outlier from the dataset.

```{r}
D_out <- d %>% filter(d$ï..Species == 'Roach' & Weight > 177.25 ) #assign the outlier data where the Species = 'Roach' and Weight> 177.25
nrow(D_out)
```

```{r}
d_new <-anti_join(d,D_out)
nrow(d_new)
```

```{r}
ggplot(d_new,aes(x=d_new$ï..Species,y=d_new$Weight,fill=d_new$ï..Species))+geom_boxplot()
```

```{r}
ggplot(d_new,aes(x=d_new$ï..Species,y=d_new$Length1,fill=d_new$ï..Species))+geom_boxplot()
```

```{r}
ggplot(d_new,aes(x=d_new$ï..Species,y=d_new$Height,fill=d_new$ï..Species))+geom_boxplot()
```






```{r}
library(caTools)
```


```{r}
#USe caret function createDataPartition to create train and test dataset (0.7: 0.3)
#LET the rows with missing NAs be the test set
split <- sample.split(d$Weight, SplitRatio = 0.7)
# Training set
train = subset(d,split=TRUE)
# Testing dataset
test = subset(d,split=FALSE)
head(train)
head(test)
dim(train)
dim(test)
summary(train)
summary(test)
```

```{r}
model=lm(d$Weight~ d$Height+d$Width+d$Length3+d$Length1, data=d)
summary(model)
```

```{r}
predicted_values <- predict(model, test[,-3], type = "response")
head(predicted_values)
```