---
title: "Time Series Forecasting with Rice Production in INDIA Dataset"
author: "Subham Das"
date: '2022-11-05'
output: word_document
---


Taking the Rice-Production Dataset

```{r}
library(readxl)
Book1 <- read_excel("D:\\DATASETS\\Rice_prod.xlsx")
Book1
```

Making it as a Time-Series Dataset.

```{r}
Rice_production=ts(Book1$`Rice production`,start=1950,end=2020,frequency = 1)
```


Time-Series plot of the Data

```{r}
library(tseries)
ts.plot(Rice_production,main='Rice production in India', xlab='Years', ylab='Production', col='darkblue')
```
We can easily see that in this dataset only Trend component is present and no Seasonal component is present.



ACF of the Dataset

```{r}
library(forecast)
Acf(Rice_production)
```
As we can see that the 12th lag is less than the 13th lag.
If seasonality is there the 13th lag should be greater than 12th lag.
So, we can say that in our dataset there is no Seasonal component.



Null Hypothesis: The series is non-stationary
Alternate Hypothesis: The series is stationary


```{r}
adf.test(Rice_production)
```

Since p value obtained 0.7145 which is greater than 5% significance level, we could conclude that the data is non-stationary.



**if the data has Trend and Seasonal component both, we can use SARIMA & Winters.**

**if we remove Seasonality component, we can use ARIMA and Holt's Exponential.**

**and if we remove Trend & Seasonality both, we can use ARMA & Simple Exponential as the data will be Stationary then.**

```{r}
length(Rice_production)
```

```{r}
Rice_production1=Rice_production[1:66]   # for in-sample forecasting
Rice_production2=ts(Rice_production1,start=1950,frequency=1)
class(Rice_production2)
```


**HOLT'S EXPONENTIAL SMOOTHING**

```{r}
library(astsa)
rice_prod_for=HoltWinters(Rice_production2,gamma = FALSE) 
rice_prod_for

#gamma is False because no seasonal component is present
```


```{r}
plot(rice_prod_for)
```


In-sample Forecasting

```{r}
library(forecast)
forecast_data=forecast(rice_prod_for,h=5)
forecast_data
```


```{r}
library(Metrics)
actual_rice_production=c(1097, 1128, 1165, 1189, 1223)
predict_rice_production=c(1080.614, 1095.139, 1109.663, 1124.188, 1138.713)
```

**Root Mean Squared Error**

```{r}
result_1=rmse(actual_rice_production,predict_rice_production)
result_1
```

**Mean Absolute Percentage Error**

```{r}
result_2=mape(actual_rice_production,predict_rice_production)
result_2
```

**Mean Absolute Scaled Error**

```{r}
result_3=mase(actual_rice_production,predict_rice_production)
result_3
```


**RESIDUAL ANALYSIS : HOLTS EXPONENTIAL SMOOTHING**

```{r}
#non-coliinearity check
res <- resid(forecast_data)
Box.test(res, type = "Ljung-Box")
```


```{r}
forecast::checkresiduals(forecast_data)
```

```{r}
# normality test
shapiro.test(res)
```


**ARIMA MODEL**

```{r}
library(forecast)
fit=auto.arima(Rice_production2)
fit
```

In-sample Forecasting

```{r}
forecast_rice_production<-forecast(fit,h=5)
forecast_rice_production
```

```{r}
library(Metrics)
actual_rice_production=c(1097, 1128, 1165, 1189, 1223)
predict_rice_production=c(1075.780, 1088.965, 1102.149, 1115.334, 1128.519)
```


**Root Mean Squared Error**

```{r}
result_1=rmse(actual_rice_production,predict_rice_production)
result_1
```

**Mean Absolute Percentage Error**

```{r}
result_2=mape(actual_rice_production,predict_rice_production)
result_2
```

**Mean Absolute Scaled Error**

```{r}
result_3=mase(actual_rice_production,predict_rice_production)
result_3
```


**RESIDUAL ANALYSIS: ARIMA**

```{r}
res <- resid(fit)
Box.test(res, type = "Ljung-Box")
```

```{r}
forecast::checkresiduals(fit)
```

```{r}
# normality test
shapiro.test(res)
```


As we can see that the RMSE of HOLT is less than ARIMA, we can say HOLT is the best model here.

**FORECASTING USING THE BEST MODEL: HOLT**

```{r}
library(astsa)
rice_prod_for1=HoltWinters(Rice_production,gamma = FALSE)
library(forecast)
forecast_data_final=forecast(rice_prod_for1,h=5)
forecast_data_final
```

```{r}
plot(forecast_data_final)
```

**Finally we have forecasted using Holts exponential smoothing for the next 5 years rice production in India.**



**Making the data Stationary by using Method of Differencing.**

```{r}
diff_Rice_production=diff(Rice_production)
```

```{r}
ts.plot(diff_Rice_production,main='Rice production in India', xlab='Years', ylab='Production', col='darkblue')
```


Null Hypothesis: The series is non-stationary
Alternate Hypothesis: The series is stationary


```{r}
adf.test(diff_Rice_production)
```

Since p value obtained 0.01 which is less than 5% significance level, we could conclude that the data is stationary.